### 2024-09-27 paper/paper.qmd
Minor fixes

---
title: "Torontonian Hate Crime Trends (temp)"
subtitle: "An Analysis of Race Bias in Hate Crime from 2018 to 2023"
author: 
  - Karen Riani
thanks: "Code and data are available at: https://github.com/karenrni/Torontonian-Hate-Crime-Trends"
date: today
date-format: long
abstract: "The abstract answers: 1) what was done, 2) what was found, and 3) why this matters (all at a high level). Likely four sentences. Abstract must make clear what we learn about the world because of this paper."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(knitr)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction

In recent years, there has been a noticeable increase in hate crimes targeting individuals of East and South Asian descent, particularly in diverse urban areas such as Toronto. These crimes have garnered significant attention amid the global rise of anti-Asian sentiment, not only in physical attacks but also in the rise of online cyberbullying. The initial spike in hate incidents was observed during the COVID-19 pandemic, primarily affecting East Asian and Chinese communities. Furthermore, changes in immigration policies, especially those impacting international students, have led to heightened animosity towards South Asian individuals. This ongoing trend shows patterns of racial bias that highlight the need to understand the dynamics of hate crimes, not only to promote public safety but also to encourage greater awareness and solidarity within the community.

Amidst the rise in Asian hate in the month of March 2020, the Canadian government began more actively working with Asian communities to handle increased anti-Asian sentiment heightened by COVID-19. Several policies were recommended by these communities, and the Canadian government responded in May 2021 by developing a definition of anti-Asian racism, which was added to Canada’s Anti-Racism Strategy (Government of Canada, 2023).

The Canadian Race Relations Foundation (CRRF) and Public Safety Canada additionally organized several engagement initiatives and published statements for Asian communities, i.e. Asian and Sikh Heritage Month and The Bias, Sensitivity, Diversity and Identity Symposium on Islamophobia and Anti-Asian Hate. These efforts provided a platform for Asian Canadians to voice their experiences and challenges. Expert discussions further highlighted the impact of national security policies on Muslim and Asian communities amid rising hate crimes.

The 2020 policy changes allowing international students to study online while preserving work permit eligibility helped mitigate disruptions but did not address the broader rise in anti-Asian racism (Government of Canada, 2020). Social media fueled this sentiment, leading to increased hate crimes against South Asian students in public and online spaces, especially in Toronto (The Link, 2022).

Despite these policy responses, Asian communities continue to face systemic discrimination. This paper examines hate crimes against Asian populations in Toronto from 2018 to 2023, with a focus on how socio-political developments, like the pandemic and policy shifts, have shaped these trends. While acknowledging broader racial disparities affecting Black communities, this paper narrows in on anti-Asian bias, aiming to inform more targeted interventions and support strategies for affected groups.

The following structure of this paper is as follows: Section 2 presents the data sources and methodology, followed by Section 3, which discusses the results and their implications. Section 4 concludes with a discussion of the limitations of the study and suggestions for future research.

# Data {#sec-mydatasection}

The data used in this paper is sourced from the City of Toronto's Open Data Portal, accessed via the opendatatoronto package (Gelfand 2022). The dataset, Hate Crimes Open Data (City of Toronto 2024), includes verified hate crime occurrences reported to the Toronto Police since 2018. Analysis was conducted using R (R Core Team 2023), with key packages dplyr (Wickham et al. 2023), tibble (Müller et al. 2023), tidyverse (Wickam et al. 2019), and ggplot2 (Wickham 2016). Report generation was done with knitr (Xie 2023).

## Hate Crime Data

Hate crime data provided by the City of Toronto’s Hate Crime Unit includes any reported hate crime occurrence beginning from January 2018. The features include identifiers like offence and report details (ID, year, date, time), police division and location, along with biases related to age, disability, race, ethnicity, language, religion, sexual orientation, and gender, as well as neighbourhood identifiers and whether an arrest was made.

The set includes 1350 hate crimes from January 2018 to most recently, December 2023. It is annually refreshed, with the latest occurring on September 9th, 2024. As for quality, the data has an overall score of 96% grade of gold, and above 97% completeness, usability and freshness.

```{r}
#| label: fig-Race-Bias-Fig
#| fig-cap: Race Biased Hate Crimes in Toronto
#| echo: false

#Count total hate crimes
hate_crime_counts <- data %>%
  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
  summarise(count = n(), .groups = 'drop')

## Plotting ##

# Find the top 4 races with the most hate crimes per year
top_per_year <- hate_crime_counts %>%
  group_by(OCCURRENCE_YEAR) %>%
  top_n(4, wt = count)

# Plotting and detailing
ggplot(top_per_year, aes(x = OCCURRENCE_YEAR, y = count, fill = RACE_BIAS)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Top 4 Hate Crimes by Race Bias (2018-2023)",
       x = "Year",
       y = "Number of Hate Crimes",
       fill = "Race Bias") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8))      
```

Seen via (@fig-Asian-Bias-Fig).

```{r}
#| label: fig-Asian-Bias-Fig
#| fig-cap: Asian Biased Hate Crimes in Toronto
#| echo: false

# To filter for Asian races
asian_races <- c("East/Southeast Asian", 
                 "South Asian", 
                 "Black, East/Southeast Asian", 
                 "Black, South Asian", 
                 "East/Southeast Asian, South Asian, Black", 
                 "White, East/Southeast Asian",
                 "Indigenous, East/Southeast Asian")

# Filter for Asian-related data
asian_data <- data %>%
  filter(RACE_BIAS %in% asian_races)

# Aggregate data by year and race bias
aggregated_data <- asian_data %>%
  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
  summarise(incident_count = n(), .groups = 'drop')

# Create a plot
ggplot(aggregated_data, aes(x = OCCURRENCE_YEAR, y = incident_count, fill = RACE_BIAS)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Hate Crimes Involving Asian Races in Toronto (2018-2024)",
       x = "Year",
       y = "Number of Incidents",
       fill = "Race Bias") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.text = element_text(size = 7))

```

```{r}
#| label: fig-Asian-Bias-Summary
#| fig-cap: Summary of Asian Biased Hate Crimes in Toronto
#| echo: false

# Generate summary statistics for Asian bias data 
summary <- asian_data %>%
  group_by(OCCURRENCE_YEAR) %>%
  summarise(
    `Incident Count` = n(),
    `Race Biases` = n_distinct(RACE_BIAS),
    `Location Types` = n_distinct(LOCATION_TYPE),
    `Offence Groups` = n_distinct(OFFENCE_GROUP)
  ) %>%
  rename(`Year` = OCCURRENCE_YEAR)

# Display the summary as a table using kable
kable(summary, caption = "Summary of Asian Biased Hate Crimes in Toronto")

```

```{r}
#| label: fig-East-Asn-Summary
#| fig-cap: Summary of increase in hate crimes against East/Southeast Asians
#| echo: false

# Group by OCCURRENCE_YEAR and RACE_BIAS to summarize the counts
Easian_crimes <- data %>%
  filter(RACE_BIAS %in% c("East/Southeast Asian")) %>%
  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
  summarise(count = n(), .groups = 'drop')

# Calculate year-on-year increase
Easian_crimes <- Easian_crimes %>%
  arrange(RACE_BIAS, OCCURRENCE_YEAR) %>%
  group_by(RACE_BIAS) %>%
  mutate(increase = (count - lag(count)) / lag(count) * 100)  # Percentage increase

# Display the result as a table
Easian_crimes %>%
  kable(caption = "Year-on-Year Increase in Hate Crimes Against Asians")

```

```{r}
#| label: fig-South-Asn-Summary
#| fig-cap: Summary of increase in hate crimes against South Asians
#| echo: false

# Group by OCCURRENCE_YEAR and RACE_BIAS to summarize the counts
Sasian_crimes <- data %>%
  filter(RACE_BIAS %in% c("South Asian")) %>%
  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
  summarise(count = n(), .groups = 'drop')

# Calculate year-on-year increase
Sasian_crimes <- Sasian_crimes %>%
  arrange(RACE_BIAS, OCCURRENCE_YEAR) %>%
  group_by(RACE_BIAS) %>%
  mutate(increase = (count - lag(count)) / lag(count) * 100)  # Percentage increase

# Display the result as a table
Sasian_crimes %>%
  kable(caption = "Year-on-Year Increase in Hate Crimes Against Asians")

```
### Raw Data
source, structure,

### Cleaned Data
recategorization of variables.

### Basic Summary Statistics of the Data

### Discussion of Data Selection
Discuss the rationale behind the selection of the data used for analysis and any criteria applied.

## Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

# Limitations and Next Steps

- accuracy and reliability of reports
- dataset includes only all verified Hate Crime occurrences investigated by the Hate Crime Unit by 
- only those confirmed and reported (date since 2018) 
- does not include occurrences that have been deemed unfounded or classified as hate incidents
- includes interaction between multi-biased hate crimes, where motivation lacks clarity 
  - i.e. Age, Mental or Physical Disability, Race, Ethnicity, Language, Religion, Sexual Orientation, Gender and Other Similar Factor
- should contine to analyse with multi level bias crimes in mind
- data is provided at the offence and/or occurrence level-- one occurrence may have multi-bias categories associated to the victim used to categorize the hate crime
- may overlook qualitative aspects of hate crimes, such as the socio-cultural context of incidents

\newpage

# References

City of Toronto. 2024. Hate Crimes Open Data. Published by Policy, Planning, Finance & Administration. Retrieved from https://www.toronto.ca/open-data-catalogue/hate-crimes-open-data.

Gelfand, Sharla. 2022. Opendatatoronto: Access the City of Toronto Open Data Portal. Retrieved from https://CRAN.R-project.org/package=opendatatoronto.

Government of Canada. 2020. Measures to Support International Students During the COVID-19 Pandemic. Immigration, Refugees and Citizenship Canada. Retrieved from https://www.canada.ca/en/immigration-refugees-citizenship/services/coronavirus-covid19.html.

Government of Canada. 2023. Annual Report on the Operation of the Canadian Multiculturalism Act 2021-2022. Retrieved from https://www.canada.ca/en/canadian-heritage/corporate/publications/plans-reports/annual-report-canadian-multiculturalism-act-2021-2022.html.

Hadley Wickham, Lionel Henry, Romain François. 2023. Dplyr: A Grammar of Data Manipulation. Retrieved from https://CRAN.R-project.org/package=dplyr.

Kirill Müller, Romain Francois, Hadley Wickham. 2023. Tibble: Simple Data Frames. Retrieved from https://CRAN.R-project.org/package=tibble.

The Link. 2022. Canada Has a Rampant South Asian Racism Problem. The Link Newspaper. Retrieved from https://thelinknewspaper.ca/article/canada-has-a-rampant-south-asian-racism-problem.

R Core Team. 2023. R: A Language and Environment for Statistical Computing. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from https://www.R-project.org/.

Wickham, Hadley. 2016. Ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York. Retrieved from https://ggplot2.tidyverse.org.

Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D’Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019. “Welcome to the Tidyverse.” Journal of Open Source Software 4 (43): 1686. Retrieved from https://doi.org/10.21105/joss.01686.

Xie, Yihui. 2023. Knitr: A General-Purpose Package for Dynamic Report Generation in R. Retrieved from https://yihui.org/knitr/.