### 2024-09-28 paper/paper.qmd
finalized paper

---
title: "Asian Biased Hate Crime Trends in Toronto"
subtitle: "An Analysis of the increased Asian Race Bias in Hate Crimes from 2018 to 2023"
author: 
  - Karen Riani
thanks: "Code and data are available at: https://github.com/karenrni/Torontonian-Hate-Crime-Trends"
date: today
date-format: long
abstract: "In this study, we analyzed hate crime trends in Toronto from 2018 to 2023, focusing on racial bias against East, Southeast, and South Asian communities. Most incidents occurred in public settings, such as businesses or on the street, with offences primarily involving assault, threats, or mischief. Results aligned with major global events and found that hate crimes against East/Southeast Asians surged during the COVID-19 pandemic, while incidents targeting South Asian communities steadily increased beginning from COVID-19. These findings highlight the role of public spaces in hate crime incidents and emphasize the need for targeted interventions to address both short-term spikes and persistent racial bias in Toronto."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(knitr)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction

In recent years, there has been a noticeable increase in hate crimes targeting individuals of East and South Asian descent, particularly in diverse urban areas such as Toronto. These crimes have garnered significant attention amid the global rise of anti-Asian sentiment, not only in physical attacks but also in the rise of online cyberbullying. The initial spike in hate incidents was observed during the COVID-19 pandemic, primarily affecting East Asian and Chinese communities. Furthermore, changes in immigration policies, especially those impacting international students, have led to heightened animosity towards South Asian individuals. This ongoing trend shows patterns of racial bias that highlight the need to understand the dynamics of hate crimes, not only to promote public safety but also to encourage greater awareness and solidarity within the community.

Amidst the rise in Asian hate in the month of March 2020, the Canadian government began more actively working with Asian communities to handle increased anti-Asian sentiment heightened by COVID-19. Several policies were recommended by these communities, and the Canadian government responded in May 2021 by developing a definition of anti-Asian racism, which was added to Canada’s Anti-Racism Strategy (@gov_canada_2023_multiculturalism).

The Canadian Race Relations Foundation (CRRF) and Public Safety Canada additionally organized several engagement initiatives and published statements for Asian communities, i.e. Asian and Sikh Heritage Month and The Bias, Sensitivity, Diversity and Identity Symposium on Islamophobia and Anti-Asian Hate. These efforts provided a platform for Asian Canadians to voice their experiences and challenges. Expert discussions further highlighted the impact of national security policies on Muslim and Asian communities amid rising hate crimes.

The 2020 policy changes allowing international students to study online while preserving work permit eligibility helped mitigate disruptions but did not address the broader rise in anti-Asian racism (@gov_canada_2020_students). Social media fueled this sentiment, leading to increased hate crimes against South Asian students in public and online spaces, especially in Toronto (@the_link_2022).

Despite these policy responses, Asian communities continue to face systemic discrimination. This paper examines hate crimes against Asian populations in Toronto from 2018 to 2023, with a focus on how socio-political developments, like the pandemic and policy shifts, have shaped these trends. While acknowledging broader racial disparities affecting Black communities, this paper narrows in on anti-Asian bias, aiming to inform more targeted interventions and support strategies for affected groups.

The following structure of this paper is as follows: Section @sec-mydatasection presents the data sources and methodology, followed by Section 3, which discusses the results and their implications. Section 4 concludes with a discussion of the limitations of the study and suggestions for future research.

# Data {#sec-mydatasection}

The data used in this paper is sourced from the City of Toronto's Open Data Portal, accessed via the opendatatoronto package (@opendatatoronto). The dataset, Hate Crimes Open Data (@toronto_hatecrimes_2024), includes verified hate crime occurrences reported to the Toronto Police since 2018. Analysis was conducted using R (@r_core_team_2023), with key packages such as dplyr (@dplyr_2023), tibble (@tibble_2023), tidyverse (@tidyverse_welcome_2019), and ggplot2 (@ggplot2_wickham_2016). Report generation was done with knitr (@knitr_xie_2023).

## Hate Crime Data

Hate crime data provided by the City of Toronto’s Hate Crime Unit includes any reported hate crime occurrence beginning from January 2018. The features include identifiers like offence and report details (ID, year, date, time), police division and location, along with biases related to age, disability, race, ethnicity, language, religion, sexual orientation, and gender, as well as neighbourhood identifiers and whether an arrest was made. These biases are defined as “crimes committed on the basis of the Suspect’s perception of the Victim’s race”, (@toronto_hatecrimes_2024).

The set includes 1350 hate crimes from January 2018 to most recently, December 2023. It is annually refreshed, with the latest occurring on September 9th, 2024. As for quality, the data has an overall score of 96% grade of gold, and above 97% completeness, usability and freshness. The initial set of data included all races and mixes, with the column “Other race not mentioned” renamed to “Unknown” and all crimes not mainly associated with race bias removed– any row containing “None” under RACE_BIAS.

The columns included in the analysis are Primary Offences (Regrouped), Race Bias and Occurance Year. Due to the analysis’ focus on Asian Bias hate crimes, races of only East, Southeast and South Asian descent were included. Furthermore, primary offences were regrouped into more general overarching categories including, Assault-Related, Threats and Harassment, Mischief & Property Damage, Hate-Related and Others. Only the occurrence year was included as opposed to full dates, and reported dates were used to focus on high-level trends. Additionally, the dataset did not contain information on other socio-economic factors related to location, such as mean income levels and ethnic demographics, thus locations were excluded. Furthermore, location type was recategorized into more general categories including public, private, government, casual and other places.

# Results

## Overview of Hate Crimes by Race
We are primarily interested in hate crimes targeting Asian races, as this aligns with our focus on anti-Asian sentiment. However, it is also important to examine hate crimes across all racial categories to provide a comprehensive understanding of the trends. In @fig-Race-Bias-Fig, the graph hows a consistent increase in hate crimes targeting the Black community. Notably, hate crimes against Black individuals have been significantly higher than those targeting other racial groups throughout the observed period.

```{r}
#| label: fig-Race-Bias-Fig
#| fig-cap: Race Biased Hate Crimes in Toronto
#| echo: false

#Count total hate crimes
hate_crime_counts <- data %>%
  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
  summarise(count = n(), .groups = 'drop')

## Plotting ##

# Find the top 4 races with the most hate crimes per year
top_per_year <- hate_crime_counts %>%
  group_by(OCCURRENCE_YEAR) %>%
  top_n(4, wt = count)

# Plotting and detailing
ggplot(top_per_year, aes(x = OCCURRENCE_YEAR, y = count, fill = RACE_BIAS)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Top 4 Hate Crimes by Race Bias (2018-2023)",
       x = "Year",
       y = "Number of Hate Crimes",
       fill = "Race Bias") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8))      
```

## Hate Crimes Against Asian Communities

@fig-Asian-Bias-Fig highlights the trend in hate crimes targeting all Asian racial groups, with a year-by-year breakdown. This visualization demonstrates a rise in hate crimes against Asian communities, especially during the years corresponding to the COVID-19 pandemic. Although the overall trend for Asian hate crimes has fluctuated, the data shows an alarming increase starting in 2020.

```{r}
#| label: fig-Asian-Bias-Fig
#| fig-cap: Asian Biased Hate Crimes in Toronto
#| echo: false

# To filter for Asian races
asian_races <- c("East/Southeast Asian", 
                 "South Asian", 
                 "Black, East/Southeast Asian", 
                 "Black, South Asian", 
                 "East/Southeast Asian, South Asian, Black", 
                 "White, East/Southeast Asian",
                 "Indigenous, East/Southeast Asian")

# Filter for Asian-related data
asian_data <- data %>%
  filter(RACE_BIAS %in% asian_races)

# Aggregate data by year and race bias
aggregated_data <- asian_data %>%
  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
  summarise(incident_count = n(), .groups = 'drop')

# Create a plot
ggplot(aggregated_data, aes(x = OCCURRENCE_YEAR, y = incident_count, fill = RACE_BIAS)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Hate Crimes Involving Asian Races in Toronto (2018-2024)",
       x = "Year",
       y = "Number of Incidents",
       fill = "Race Bias") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.text = element_text(size = 7))

```

## Summary Statistics: East/Southeast Asian and South Asian Communities

@tbl-Asian-Bias-Summary presents summary statistics for hate crimes specifically targeting Asian racial groups. @tbl-East-Asn-Summary and @tbl-South-Asn-Summary show the increase in the 2 major Asian race groups being East/Southeast Asians and South Asians, respectively. The data reveals a steady increase in incidents against South Asians in 2021, with a notable spike in crimes against East Asians. The increase in Asian biased crimes also show an increase in different kinds of offences and location of hate crime occurrences.

```{r}
#| label: tbl-Asian-Bias-Summary
#| table-cap: Summary of Asian Biased Hate Crimes in Toronto
#| echo: false

# Generate summary statistics for Asian bias data 
summary <- asian_data %>%
  group_by(OCCURRENCE_YEAR) %>%
  summarise(
    `Incident Count` = n(),
    `Race Biases` = n_distinct(RACE_BIAS),
    `Location Types` = n_distinct(LOCATION_TYPE),
    `Offence Groups` = n_distinct(OFFENCE_GROUP)
  ) %>%
  rename(`Year` = OCCURRENCE_YEAR)

# Display the summary as a table using kable
kable(summary, caption = "Summary of Asian Biased Hate Crimes in Toronto")

```

```{r}
#| label: tbl-East-Asn-Summary
#| table-cap: Summary of increase in hate crimes against East/Southeast Asians
#| echo: false

# Group by OCCURRENCE_YEAR and RACE_BIAS to summarize the counts
Easian_crimes <- data %>%
  filter(RACE_BIAS %in% c("East/Southeast Asian")) %>%
  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
  summarise(count = n(), .groups = 'drop')

# Calculate year-on-year increase
Easian_crimes <- Easian_crimes %>%
  arrange(RACE_BIAS, OCCURRENCE_YEAR) %>%
  group_by(RACE_BIAS) %>%
  mutate(increase = (count - lag(count)) / lag(count) * 100)  # Percentage increase

# Display the result as a table
Easian_crimes %>%
  kable(caption = "Year-on-Year Increase in Hate Crimes Against East/ Southeast Asians")

```

```{r}
#| label: tbl-South-Asn-Summary
#| table-cap: Summary of increase in hate crimes against South Asians
#| echo: false

# Group by OCCURRENCE_YEAR and RACE_BIAS to summarize the counts
Sasian_crimes <- data %>%
  filter(RACE_BIAS %in% c("South Asian")) %>%
  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
  summarise(count = n(), .groups = 'drop')

# Calculate year-on-year increase
Sasian_crimes <- Sasian_crimes %>%
  arrange(RACE_BIAS, OCCURRENCE_YEAR) %>%
  group_by(RACE_BIAS) %>%
  mutate(increase = (count - lag(count)) / lag(count) * 100)  # Percentage increase

# Display the result as a table
Sasian_crimes %>%
  kable(caption = "Year-on-Year Increase in Hate Crimes Against South Asians")

```

## Offence and location trends in Asian Biased Hate Crimes
@fig-Asian-Offence-Types and @fig-Asian-Location-Types show spikes in assault and public based hate crimes in 2021 towards East/Southeast and South Asians, including racially mixed Asians.

```{r}
#| label: fig-Asian-Offence-Types
#| fig-cap: Type of Offences in Asian Biased Hate Crimes
#| echo: false

# Group by OCCURRENCE_YEAR and OFFENCE_GROUP for Asian hate crimes
asian_offense_data <- asian_data %>%
  group_by(OCCURRENCE_YEAR, OFFENCE_GROUP) %>%
  summarise(incident_count = n(), .groups = 'drop')

# Plotting offense types
ggplot(asian_offense_data, aes(x = OCCURRENCE_YEAR, y = incident_count, fill = OFFENCE_GROUP)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Type of Offences in Asian Biased Hate Crimes (2018-2023)",
       x = "Year",
       y = "Number of Incidents",
       fill = "Offence Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.text = element_text(size = 7))
```

```{r}
#| label: fig-Asian-Location-Types
#| fig-cap: Location Types in Asian Biased Hate Crimes
#| echo: false

# Group by OCCURRENCE_YEAR and LOCATION_TYPE for Asian hate crimes
asian_location_data <- asian_data %>%
  group_by(OCCURRENCE_YEAR, LOCATION_GROUP) %>%
  summarise(incident_count = n(), .groups = 'drop')

# Plotting location types
ggplot(asian_location_data, aes(x = OCCURRENCE_YEAR, y = incident_count, fill = LOCATION_GROUP)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Location Types in Asian Biased Hate Crimes (2018-2023)",
       x = "Year",
       y = "Number of Incidents",
       fill = "Location Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.text = element_text(size = 7))

```

# Discussion

Based on the above figures, there is evidence that the surge in hate crimes against East/Southeast and South Asians during 2021 correlates strongly with the social and political context surrounding the COVID-19 pandemic. Specifically a spike of over 1000% in hate crimes motivated by anti-East/Southeast Asian sentiment in @tbl-East-Asn-Summary. As the virus's origins were linked to Asia, widespread misinformation and xenophobia fueled anti-Asian sentiment. The sharp rise in hate crimes during this period is indicative of a broader global trend of increased racial targeting and discrimination, but concentrated within denser, diverse urban populations. While 2022 and 2023 saw a reduction in incidents, the numbers remain higher than pre-pandemic levels, suggesting that the effects of this wave of racism may have long-term impacts. Notably in @fig-Asian-Offence-Types and @fig-Asian-Location-Types, public settings such as businesses and streets became hotspots for these crimes, with assaults making up a significant portion of the incidents. In 2021, public assaults against Asians peaked and remained higher than pre-pandemic levels even into 2022 and 2023. While there was a reduction in incidents in these later years, the overall numbers still surpass those observed before the pandemic, suggesting long-lasting impacts of this wave of racism.

In contrast, the steady rise since 2021 in @tbl-South-Asn-Summary in hate crimes against South Asians points to a broader, underlying trend of racial bias that extends beyond the pandemic. Anti-immigrant sentiment, exacerbated by policies affecting international students and immigration more generally, likely played a role in these trends. Despite government interventions aimed at supporting immigrant communities during the pandemic (@gov_canada_2020_students), the persistent hostility faced by South Asians reflects the ongoing challenges of systemic racism in Canadian society.

Meanwhile, the consistent levels of anti-Black hate crimes highlight ongoing systemic racism, highlighted in @fig-Race-Bias-Fig. Although hate crimes against Asians have decreased since their peak, the persistence of incidents reflects the long-term impact of these biases, necessitating continued policy focus on combating racial discrimination.

Both the rise and persistence of these hate crimes underscore the importance of targeted policy responses. While public health measures and immigration policies were necessary during the pandemic, they were insufficient in addressing the racial tensions that surfaced. There is a clear need for long-term strategies aimed at combating racial bias and xenophobia, especially as it pertains to both East/Southeast and South Asian communities in Canada.

# Limitations and Next Steps

There are several limitations to this analysis that should be considered when interpreting the results. The accuracy and reliability of reports may be affected by underreporting or misclassification, as the dataset includes only reported and verified hate crimes by the Hate crime Unit. Many incidents may go unreported or are excluded occurrences classified as unfounded or those identified as hate incidents. As a result, the data may not fully reflect the actual or full range of hate crimes occurring in the GTA community. The limited focus may omit significant portions of incidents not classified as hate crimes which still reflect underlying societal tensions.

Some incidents in involve multi-bias hate crimes, where multiple motivations such as race, age, disability, and religion can overlap. This can make it difficult to isolate patterns of specific bias categories, as one crime may reflect more than one motivation. Understanding the full nature of these crimes becomes more complex when multiple biases are at play.

The dataset lacks important qualitative insights into the socio-cultural context of hate crimes. Factors such as mean area income, community dynamics, media influence, and social tensions that may drive hate crimes are not reflected in the dataset, which limits a full understanding of the underlying causes and conditions of these incidents.

In future research, there should be continued focus on multi-bias hate crimes in greater detail, perhaps disaggregating incidents with multiple biases to better understand intersections between race and other categories like age and religion. This will help clarify how different forms of discrimination intersect in hate crime occurrences. Additionally, it would be beneficial to incorporate qualitative data into future studies. Interviews with victims, community feedback, and analysis of media or social media trends could provide richer insights into the social and cultural factors that drive hate crimes. It may also be valuable to collaborate with community organizations to gather additional data on underreported hate crimes. Partnering with other groups, especially Asian based communities, may help fill gaps in official reports and ensure that marginalized voices are included.

Quantitative additions regarding income levels may also help in understanding the effects of major socio-economic factors on hate crime. Expanding the dataset to include a longer time frame beyond 2018 and considering incidents classified as hate-related but not fully investigated as hate crimes could provide a broader perspective on trends over time. Such an expansion would allow for more comprehensive insights into the long-term patterns of hate incidents in Toronto.

Furthermore, continued exploration of offense and location trends can be vital Analyzing the types of offenses and the locations where hate crimes against specific groups, particularly Black, East/Southeast Asians and South Asians, occur can help identify hotspots and recurring incidents. This information could be used to inform targeted interventions and policy efforts aimed at reducing hate crime occurrences in these areas.

By addressing these limitations and building on the current analysis, future research can provide a deeper and more complex understanding of hate crimes and motivations in Toronto and support more effective strategies for combating them.

\newpage

# References